
---
title: "Output"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Giving is one way to get involve with the community. Statistics Canada has been collecting data on giving, volunteering,and participating since 1997. In this paper, we focused on exploring the giving aspect, especially the reasons for giving, the decisions on giving and the behavioral patterns on giving. We found that about 68% of respondents read information provided by charity before making a decision. The reasons for giving all have about 37%. The analysis and monitoring of these social trends give a reflection of the living conditions of the Canadian public and has significant implications for public policy and program decisions."
output:
  bookdown::pdf_document2
toc: FALSE
#bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

In a highly materialistic society, personal interest and profit-making have become core values. We are seeing a decline in charitable giving in recent years. Giving reflects the involvement with the community and has an impact on it. 
Statistics Canada, along with other bodies, has collected data on giving, volunteering and participating. The main objectives of the GSS GVP are to collect data on a regular basis for analysis of societal trends and monitoring any changes in the social behavior of the Canadian public. This would provide vital information to governmental bodies and have implications for policy-making and program decisions.
 
 
Indeed, about 10 billion dollars are donated by Canadians. However, this number is on the decline.

This paper focused on the giving aspect. We conducted a descriptive analysis of the giving aspect. Our aim to understand the behavioral pattern on giving of the Canadian public.

We obtained data through the University of Toronto



The 2018 GSS GVP measures changes in society related to living conditions and well-being. It provides data to important stakeholders such as the Canadian System of National Accounts.

# Data

We obtained the 2018 GSS survey data on Giving, Volunteering and Participating through the University of Toronto library. 

Sample frame: responders eligible for interview has to be 15 or above from Canada's ten provinces. The 2018 GSS GVP was conducted from September 4th to December 28th 2018. In previous GVP surveys, only paper questionnaire and telephone were the methods of data collection. However, for the 2018 iteration, an online option was made available.

The ones collecting the data are aware that the data may face sampling and non-sampling errors.

It is the seventh time the GVP was run. It was conducted by Statistics Canada. 2018 was the first time that online questionnaire can be completed. 



## Survey Methodology

How respondents were selected?
It applied the age-order selection method. It is important to note that all respondents were not contacted the same way; some were contacted through mail, others by telephone.

Changes to survey questionnaire
content was updated from 2004 to 2018 using experience from prior surveys, notably clarity of questions and consultations with stakeholders from charity organizations. 

Another important feature of the 2018 GSS GVP is that not all respondents have the same number of questions. For example, only respondents who donated at least 10 dollars were  asked questions about decisions and reasons for giving. And those who donated more than 1150 dollars did not have to answer why they did not choose to donate more.

For the online questionnaire, a pilot test was conducted online and by telephone from February 2018 to March 30 2018.

sample frame
it has not been consistent over the years. In 2004, 2007 and 2010, random digit dialling was used. From 2013, the sample frame included "cell phone only" households, which is a population not covered by RDD. 

Internet option - adapt to technology and demands on Canadians'time. - flexibility and convenience

target population
all persons who are at least 15 years old with the exceptions of residents of the Yukon, Northwest Territories and Nunavut as well as full-time residents of institutions

proceeded with stratification
stratification sampling with the provinces divided into strata and many of the Census Metropolitan Areas (CMA) considered as separate strata. totaling 27 stata

frame
2 parts - lists of telephone numbers in use (both landline and cellular) available to Statistics Canada
- address register for mail code for online option

for each stratum, a siimple random sample without replacement was selected.

rejective sampling - 'rejecting' a portion of population with a given probability to allow more time and effort to be spent trying to find the population of interest, volunteers

target sample size was 20000 but actual number is 16149
minimum sample size for each stratum was determined beforehand to ensure acceptable sample variability

majority of interviewers had previous experience with gss or csgvp

- explain importance of survey and encourage participation

response rate was 41.9%
less than previous gss but not comparable



## Discussion and Results




strength
- large coverage with detailed/ specific questions
- well documented
- 

weaknesses
- too specific, might take too long to respond to all
- some data obtained from other sources like income from CRA 2017 tax filing, consent for survey?
- first time online version, may not be consistent with previous versions
- data quality with online questionnaire
- 

```{r library, echo=FALSE, message=FALSE}
# install.packages('bookdown')
library(dplyr)
library(tidyverse) 
library(janitor)
library(ggplot2)
library(knitr)
library(kableExtra)
library(bookdown)
```

```{r read_data, echo=FALSE, message=FALSE, warning=FALSE}
gss<- read_csv('data1.csv', show_col_types = FALSE)
raw_data <- read_csv('gssv.csv')
```


### Age

```{r age, echo=FALSE, fig.cap='Distribution of age-groups and gender in the 2018 GSS survey'}
age <-
  gss %>%
  mutate(age_group = 
           case_when(age == "15 to 24 years" ~ "15-24",
                     age == "25 to 34 years" ~ "25-34",
                     age == "35 to 44 years" ~ "35-44",
                     age == "45 to 54 years" ~ "45-54",
                     age == "55 to 64 years" ~ "55-64",
                     age == "65 to 74 years" ~ "65-74",
                     age == "75 years and over" ~ ">=75"
                     ),
         age_group = factor(age_group,
                            levels = c("15-24",
                                       "25-34",
                                       "35-44",
                                       "45-54",
                                       "55-64",
                                       "65-74",
                                       ">=75"
                                       )
                            )
         )
age %>%
  ggplot(mapping=aes(x=age_group, fill=gender)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(x='Age-group of respondent', y='Number of respondents', 
       title='Distribution of age-groups and gender in the 2018 GSS survey',
       fill='Gender')
```

The age distribution of survey respondents is shown in Figure \@ref(fig:age). Among the selected household, one member was chosen to represent the household. It shows that it was more likely that the female in their middle ages was selected to complete the questionnaire. 

### Marital Status

```{r maritalstatus, echo=FALSE, fig.cap='Marital Status'}
income <-
  gss %>%
  mutate(fam_income_quintile = 
           case_when(fam_income_quintile == "Highest quintile" ~ "Highest",
                     fam_income_quintile == "Second quintile" ~ "Second",
                     fam_income_quintile == "Third quintile" ~ "Third",
                     fam_income_quintile == "Fourth quintile" ~ "Fourth",
                     fam_income_quintile == "Lowest quintile" ~ "Lowest",
                     ),
         fam_income_quintile = factor(fam_income_quintile,
                                      levels = c('Highest',
                                                 'Second',
                                                 'Third',
                                                 'Fourth',
                                                 'Lowest')
         )
  )

income %>%
  ggplot(mapping=aes(x=marital_status))+
  geom_bar() + 
  theme_minimal() +
  labs(x='') +
  theme(axis.text.x = element_text(angle=15))
```

```{r maritalandincome, echo=FALSE, fig.cap='Marital Status and Family Income Quintile'}

income %>%
  ggplot(mapping=aes(x=fam_income_quintile, fill=fam_income_quintile)) +
  geom_bar() +
  theme_minimal() +
  labs(x='') +
  theme(axis.text.x = element_text(angle=15), legend.position='none') +
  facet_wrap(vars(marital_status),
             dir = "v",
             scales = "free")
```

Marital status is important in tracking the evolution of social attitudes. Figure \@ref(fig:maritalstatus) indicates that the most common family structure in Canada is married couples. When the marital status is combined with the family income, it can be used to indicate the socio-economic status of various types of families. Figure \@ref(fig:maritalandincome) tells that the distribution of the family income quintile is similar between the married couples and the common-law couples, and the remaining tends to have the similar distribution.

### Giving

```{r donation, echo=FALSE, fig.cap='Whether they have made the donation in the past 12 months'}
giving <-
  gss %>%
    count(giving_flag)

giving %>%
  ggplot(mapping=aes(x=giving_flag, y=n, fill=giving_flag)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(label=paste(round(n/sum(n),4)*100, '%')), size=4, position=position_stack(vjust = 0.5)) +
  theme_minimal() +
  labs(x='',y='count') +
  theme(legend.position='none')
```

Figure \@ref(fig:donation) shows that 75.26% of the respondents have answered that they have made any financial donations in the past 12 months.

### Awareness

```{r awareness, echo=FALSE, fig.cap="Awareness of the charity organizations"}
dg <- gss %>% 
  select(starts_with('decision_on_giving_'))

dg_awareness<-
  dg %>%
    select(decision_on_giving_know_how_to_verify, decision_on_giving_org_monitoring)

awareness_lst<-c('Know how to verify whether an organization is a registered charity',
                 'Monitor how charities in Canada use their donations')

df_awareness<-data.frame('Items'=c(''),
                         'No'=c(''),
                         'Yes'=c(''),
                         'NA'=c(''))
i <- 1
for (r in colnames(dg_awareness)){
  counted<-
    dg_awareness %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  
  df_awareness<-rbind(df_awareness, c(awareness_lst[i], counted$proportion))
  i<-i+1
}

df_awareness<-df_awareness[-c(1),]

df_awareness %>% 
  as_tibble() %>%
  relocate(Yes, .after=Items) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Awareness of the charity organizations", format="latex") %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling()
```

Based on Table \@ref(tab:awareness), we can first observe that most people do not know how to check if a charity is registered in Canada. A "registered charity" refers to a charitable organization that is registered with the Canada Revenue Agency (@CRA). The donation made to the registered charity is exempted from paying income tax, and the charity can issue tax receipts for donations it receives (@CRA). Next, most charities openly share their information on how the donations are being used for its accountability and transparency (@charity). However, we can tell that people do not monitor how charities use their donations.

### How to decide where to donate 

```{r dgsearch, echo=FALSE, fig.cap='Do people search before making a donation?'}
# https://statisticsglobe.com/wrap-long-axis-labels-ggplot2-plot-into-multiple-lines-r
dg <-
  dg %>% 
  mutate(decision_on_giving_search =
           recode(
             decision_on_giving_search,
             'Yes'='Yes',
             'No'='No',
             'Sometimes'='Yes',
             'You never consider donating to a charity that you have no...' = 'Never considered donating'))

dg %>%
  ggplot(aes(x=decision_on_giving_search, fill=decision_on_giving_search)) +
  geom_bar() +
  theme_minimal() +
  labs(x='') + 
  theme(legend.position='none') +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))
```

Next, we observed whether people do a research on the organizations before making donations, and if they do, what methods are being used to search for the charity. Figure \@ref(fig:dgsearch) shows that if the respondent have considered making donations, then they usually search for information on the charity before giving. 

```{r search, echo=FALSE, fig.cap='What method do people use in searching a new charity?'}
dg_search_detail <-
  dg %>%
  filter(decision_on_giving_search=="Yes" | decision_on_giving_search=="Sometimes") %>%
  select(starts_with('decision_on_giving_search'))

dg_search_detail <-
  dg_search_detail %>%
  select(-decision_on_giving_search,
         -decision_on_giving_search_consult_other_source)

dg_search_df<-data.frame('Method'=c(''),
                         'No'=c(''),
                         'Yes'=c(''),
                         'NA'=c(''))

dg_search_lst<-c('Read information provided by the charity',
                 'Contact the charity directly',
                 'Consult the CRA website',
                 'Consult the website of a non-profit or private organization',
                 'A general online search',
                 'Ask someone',
                 'Other'
)

i <- 1
for (r in colnames(dg_search_detail)){
  counted<-
    dg_search_detail %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  
  dg_search_df<-rbind(dg_search_df, c(dg_search_lst[i], counted$proportion))
  i<-i+1
}

dg_search_df<-dg_search_df[-c(1),]

dg_search_df %>% 
  as_tibble() %>%
  mutate(Yes = as.numeric(Yes),
         No = as.numeric(No),
         NA. = as.numeric(NA.)) %>%
  arrange(desc(Yes)) %>%
  relocate(Yes, .after=Method) %>%
  kable(caption = "Decision On Giving search") %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling()
```

Among those who search before they donate to a charity, Table \@ref(tab:search) shows that people are mostly likely to read information provided by the charity and do a general online search. It should be noted that even though the list of the registered charitities can be viewed from the CRA website, people did not consider consulting the CRA website before making donations. 

### Reasons for giving

```{r reasonsforgiving, echo=FALSE, fig.cap='Reasons for giving'}
# choose the data who has donated in the past 12 months
reasons_for_giving <-
  gss %>%
    select(starts_with('reas_giv_'))

rg_lst<-c('Personally affected',
          'Tax Credit',
          'Religious reason',
          'Spiritual beliefs',
          'Cause',
          'Compassion',
          'Community contribution',
          'Being asked')

rg<-data.frame('Reason'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(reasons_for_giving)){
  counted<-
    reasons_for_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  rg<-rbind(rg, c(rg_lst[i], counted$proportion))
  i<-i+1
}
rg<-rg[-c(1),]

rg %>% 
  as_tibble() %>%
  mutate(Yes = as.numeric(Yes),
         No = as.numeric(No),
         NA. = as.numeric(NA.)) %>%
  arrange(desc(Yes)) %>%
  relocate(Yes, .after=Reason) %>%
  relocate(No, .after=Yes) %>%
  kable() %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling()
```

Table \@ref(tab:reasonsforgiving) demonstrates the reasons for giving. Note that only the respondents who have made the donations in the past 12 months answered this questionnaire. Table \@ref(tab:reasonsforgiving) indicates that it varies for each people why they decide to make donations. Although it turned out to be approximately equal among every reason listed in the survey, the top 2 were spiritual beliefs and religious reasons, which imply that one's faith is one of the biggest reasons why people decide to give money to the charity.

### Reasons for not giving (more)

```{r reasonsfornotgiving, echo=FALSE, fig.cap='Reasons for not giving (more)'}
reasons_for_not_giving <-
  gss %>%
    select(starts_with("reas_not_giv")) %>%
  mutate(reas_not_giv_charity_fraud = 
           case_when(reas_not_giv_charity_fraud == "Agree" ~ "Yes",
                     reas_not_giv_charity_fraud == "Disagree" ~ "No",
                     reas_not_giv_charity_fraud == "NA" ~ "NA"),
         reas_not_giv_so_many_org = 
           case_when(reas_not_giv_so_many_org == "Agree" ~ "Yes",
                     reas_not_giv_so_many_org == "Disagree" ~ "No",
                     reas_not_giv_so_many_org == "NA" ~ "NA"))

ng_lst<-c('Already gave enough',
          'Cannot afford',
          'Not being asked',
          'Did not know where to donate',
          'Hard to find cause',
          'Volunteered instead',
          'Gave directly to people',
          'Tax credit not incentive enough',
          'Money would not be used efficient',
          'Did not like the way requested',
          'So many organizations',
          'Charity fraud'
          )

ng<-data.frame('Reason'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(reasons_for_not_giving)){
  counted<-
    reasons_for_not_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  ng<-rbind(ng, c(ng_lst[i], counted$proportion))
  i<-i+1
}
ng<-ng[-c(1),]

ng %>% 
  as_tibble() %>%
  mutate(Yes = as.numeric(Yes),
         No = as.numeric(No),
         NA. = as.numeric(NA.)) %>%
  arrange(desc(Yes)) %>%
  relocate(Yes, .after=Reason) %>%
  relocate(No, .after=Yes) %>%
  kable() %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling()
```

Next, we observed for the reasons for not giving. If the respondents have made the donations in the past 12 months, then it asked for the reasons why they did not donate more. Table \@ref(tab:reasonsfornotgiving) demonstrates that among the respondents who have donated in the past 12 months, 31.17% of them think that they have already donated enough. 
When the respondents did not donate in the past 12 months, it asked for the reasons for not giving. Table \@ref(tab:reasonsfornotgiving) shows that 42.57% of the respondents decided to volunteer instead of giving. 42.41% of the respondents answered that they did not like the way they were being requested. Additionally, it is worth mentioning that although there are so many organizations, it was not the main reason why people decided not to give money. It was that they did not know where to make donations among those organizations, and they were afraid that money would not be used efficient. Furthermore, 42.32% of the respondents said that they gave money directly to people who were in need of financial assistance.

### Other givings

```{r othergiving, echo=FALSE}
other_giving<-
  gss %>%
  select(starts_with("other_giv_"))

og_lst<-c('Food',
          'Clothings, toys, or household goods',
          'Other',
          'Directly to people'
          )

og<-data.frame('Other'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(other_giving)){
  counted<-
    other_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  og<-rbind(og, c(og_lst[i], counted$proportion))
  i<-i+1
}
og<-og[-c(1),]

og %>% 
  as_tibble() %>%
  mutate(Yes = as.numeric(Yes),
         No = as.numeric(No),
         NA. = as.numeric(NA.)) %>%
  relocate(Yes, .after=Other) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Other givings") %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling()
```

Lastly, we observed the other givings they have made besides financial donations to the charity. It can be observed from Table \@ref(tab:othergiving) that 7% of the respondents have donated food to the organizations such as food bank, and 6.72% donated clothings, toys, or household goods. In addition, 6.64% have made other givings directly to the ones who were in need of help. 6.90% made a donation to a charity in their will or through another financial product such as insurance. However, we can see that people mostly make the financial donations to the charitable or non-profitable organizations. 

```{r donationresponse, fig.cap = "Comparing the responses to different methods", echo = FALSE}
# Graph showing the counts for y/n for financial giving

# select the required variables
# rename the variables
financial_giving <- gss %>% 
  select(starts_with("fin_giving_"), caseid) %>% 
  rename("mail" = "fin_giving_mail",
         "telephone" = "fin_giving_telephone",
         "tv" = "fin_giving_tv",
         "online" = "fin_giving_online",
         "own initiative" = "fin_giving_own_initiative",
         "charity event" = "fin_giving_charity_event",
         "in mem of someone" = "fin_giving_in_mem_of_someone",
         "work" = "fin_giving_work",
         "door to door canvassing" = "fin_giving_door_to_door_canvassing",
         "shopping centre" = "fin_giving_shopping_centre",
         "place of worship" = "fin_giving_place_of_worship",
         "sponsoring" = "fin_giving_sponsoring",
         "other" = "fin_giving_other")

# tidy data
financial_giving <- financial_giving %>% 
  pivot_longer(!caseid,
               names_to = "name",
               values_to = "response")

# plot data
graph_1 <-
  financial_giving %>% 
  ggplot(aes(x = name, fill = response)) +
  geom_bar(position = "dodge") +
  labs(x = "Methods responded to",
       y = "Number of survey respondents") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(palette = "PuBu")

graph_1 + theme(legend.position = "bottom")
```

Figure \@ref(fig:donationresponse) shows the number of responses of the different methods responded to. For example, online means the number of people who made donations by responding to an online request which can be an email or social media. From the figure, it is clear that people tend to positively respond to request in shopping centre while other ways of requests received the least number of donations. However, the number of donations does not necessarily mean more amount donated. 

```{r numberdonations, fig.cap="Distribution of the total number of financial donations in the past 12 months", echo = FALSE}
# total number of financial donations
# obtain the total number of financial donations from the raw dataset
total_number_fin_donation <- raw_data %>% 
  select(gs1dntot)

# rename column with a more appropriate name
clean_total_fin_donation <- total_number_fin_donation %>% 
  rename(number_of_financial_donations = gs1dntot)

# obtain the data for those who did not skip the question
clean_total_fin_donation_1 <- 
  clean_total_fin_donation %>% 
  filter(number_of_financial_donations != 96)

# number of people who skipped this question
skipped_response <-
  clean_total_fin_donation %>% 
  filter(number_of_financial_donations == '96') %>% 
  count()

# distribution of total number of financial donations made in the past 12 months
clean_total_fin_donation_1 %>% 
  ggplot(aes(x = number_of_financial_donations)) +
  geom_histogram(bins = 40, fill = "lightblue") +
  labs(x = "Total number of financial donations",
       y = "Number of donations") +
  theme_minimal()
```
Figure \@ref(fig:numberdonations) shows the distribution of the number financial donations made by the survey respondents. It is important to note that out of the 16149 respondents, 3995 people skipped the question and this number was removed before making the graph. The most number of financial donations made was 72 but with most other respondents making between 1 and 5 donations in the last 12 months.




```{r totalamount, fig.cap="Distribution of the total amount donated in the past 12 months", echo = FALSE}
# total amount of donations
total_amount <- gss %>% 
  select(totl_amount_don12) %>% 
  rename(total = totl_amount_don12)

# data for only those who responded
total_amount_1 <- total_amount %>% 
  filter(total != 999999999.96)

# number of valid skipped responses
valid_skip_responses <- total_amount %>% 
  filter(total == 999999999.96) %>% 
  count()

# distribution of total amount of donations
total_amount_1 %>% 
  ggplot(aes(x = total)) +
  geom_histogram(bins = 30, fill = "lightblue") +
  labs(x = "Total amount of donations",
       y = "Number of donations") +
  theme_minimal()

# total amount donated
total_sum <- sum(total_amount_1$total)

g <-total_amount_1 %>% 
  filter(total > 20000)
```
Figure \@ref(fig:totalamount) illustrates how the donation amounts are distributed with 0.5 dollars as the least amount donated and $60,000 as the largest amount donated in the past 12 months. 3995 respondents chose to skip this question. A total of 9.3 million dollars have been donated by the 12154 Canadian respondents. From figure \@ref(fig:totalamount), most respondents donated amounts less than 200 dollars while only 28 out of the 12154 respondents donated more than 20000 dollars.


```{r income, fig.cap = "Comparing the number of respondents in the different income groups", echo = FALSE}
# obtain income data
# rename the variables name
income_group <- raw_data %>% 
  select(incg2, famincg2) %>% 
  rename(personal_income_group = incg2,
         family_income_group = famincg2)

# recode the variable values
income_group <- income_group %>% 
  mutate(personal_income_group = 
           recode(personal_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more"),
         family_income_group = recode(
           family_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more")
           )

# distribution of income
income_group %>% 
  ggplot(aes(x = personal_income_group)) +
  geom_bar(fill = "lightblue") +
  labs(x = "Personal Income group before taxes",
       y = "Number of respondents") +
  theme_minimal() +
  coord_flip()


```
Figure \@ref(fig:income) shows the distribution of personal income groups of the survey respondents. The data was obtained using the CRA income tax filing in 2017. We can see that most respondents were from low to middle income background. The highest proportion of respondents were found in the 25000 to 49999 income group (29.95%) and 78.23% of the respondents earned less than $75000 annually before taxes.




```{r averagedonate, echo=FALSE}
# age income 
donate_income_group <- raw_data %>% 
  select(incg2, gs1datot) %>% 
  rename(personal_income_group = incg2,
         total_amount_donated = gs1datot)

donate_income_group <- donate_income_group %>% 
  mutate(personal_income_group = 
           recode(personal_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more")
           )

# create summary table

donate_income_group %>% 
  filter(total_amount_donated != 999999999.96) %>% 
  group_by(personal_income_group) %>% 
  summarise(mean = round(mean(total_amount_donated), 2),
            sd = round(sd(total_amount_donated), 2)) %>% 
  kable(caption = "Summary statistics of total amount donated per personal income group",
        col.names = c("Personal Income group", "Average amount donated", "Standard deviation of amount donaated"),
        linesep = "",
        align = c("c", "c", "c"))

```
Table \@ref(tab:averagedonate) shows the mean of the total amount donated for each personal income group as well as its variability. We can note that in general, respondents in higher income group tends have donated more on average which is not surprising. We can also observed that the mean amount donated for the 75000 to 99999 and 100000 to 124999 income group do not differ by much ($63.19) but the variability is much larger which can explain why the difference is so small. There must be some respondents who donated much larger amount compared to what others in the same income group have donated.

\newpage

# Appendix

## Supplementary survey



# extra questionnaires
- How much do you donate proportional to your income? (I donate 10% of my income every year, etc.)

- more categories for gender, with possibly a space to specify like other, specify
- when people feel like donating
- motivation for donating, in the family
- will they consider donating in the future
- 
