
---
title: "Output"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
#bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, echo=FALSE, message=FALSE}
# install.packages('bookdown')
library(dplyr)
library(tidyverse) 
library(janitor)
library(ggplot2)
library(knitr)
library(kableExtra)
library(bookdown)
```

```{r read_data, echo=FALSE, message=FALSE, warning=FALSE}
gss<- read_csv('data1.csv', show_col_types = FALSE)
raw_data <- read_csv('gssv.csv')
```

# I. Introduction

This dataset contains core and 

```{r age, echo=FALSE}
age <-
  gss %>%
  mutate(age_group = 
           case_when(age == "15 to 24 years" ~ "15-24",
                     age == "25 to 34 years" ~ "25-34",
                     age == "35 to 44 years" ~ "35-44",
                     age == "45 to 54 years" ~ "45-54",
                     age == "55 to 64 years" ~ "55-64",
                     age == "65 to 74 years" ~ "65-74",
                     age == "75 years and over" ~ ">=75"
                     ),
         age_group = factor(age_group,
                            levels = c("15-24",
                                       "25-34",
                                       "35-44",
                                       "45-54",
                                       "55-64",
                                       "65-74",
                                       ">=75"
                                       )
                            )
         )
age %>%
  ggplot(mapping=aes(x=age_group, fill=gender)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(x='Age-group of respondent', y='Number of respondents', 
       title='Distribution of age-groups and gender in the 2018 GSS survey',
       fill='Gender')
```

```{r maritalstatus, echo=FALSE, fig.cap='Marital Status and Income'}
age %>%
  ggplot(mapping=aes(x=fam_income_quintile)) +
  geom_bar() +
  theme_minimal() +
  labs(x='', title='Marital Status') +
  theme(axis.text.x = element_text(angle=15)) +
  facet_wrap(vars(marital_status),
             dir = "v",
             scales = "free")
```

Marital status is important in tracking the evolution of social attitudes. The combination of the marital status and family income indicates the socio-economic status of various types of families. 

```{r decisionongiving, echo=FALSE, fig.cap='Decision'}

dg <- gss %>% 
  select(starts_with('decision_on_giving_'))

dg_decide_in_advance <-
  dg %>%
  group_by(decision_on_giving_decide_in_advance) %>%
  summarise(percent=100*round(n()/nrow(dg),4))

dg %>%
  ggplot(aes(x=decision_on_giving_decide_in_advance)) +
  geom_bar() +
  theme_minimal()

```

```{r awareness, echo=FALSE, fig.cap="Awareness of the charity organizations"}
dg_awareness<-
  dg %>%
    select(decision_on_giving_know_how_to_verify, decision_on_giving_org_monitoring)

awareness_lst<-c('Know how to verify whether an organization is a registered charity',
                 'Monitor how charities in Canada use their donations')

df_awareness<-data.frame('Method'=c(''),
                         'No'=c(''),
                         'Yes'=c(''),
                         'NA'=c(''))
i <- 1
for (r in colnames(dg_awareness)){
  counted<-
    dg_awareness %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  
  df_awareness<-rbind(df_awareness, c(awareness_lst[i], counted$proportion))
  i<-i+1
}

df_awareness<-df_awareness[-c(1),]

df_awareness %>% 
  as_tibble() %>%
  relocate(Yes, .after=Method) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Awareness of the charity organizations", format="latex") %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling()
```






```{r dgsearch, echo=FALSE}
dg_search <-
  dg %>% 
  mutate(decision_on_giving_search =
           recode(
             decision_on_giving_search,
             'Yes'='Yes',
             'No'='No',
             'Sometimes'='Sometimes',
             'You never consider donating to a charity that you have no...' = 'Never considered donating to a new charity')) %>%
  count(decision_on_giving_search) %>% 
  mutate(proportion = round(n/sum(n),4)*100)

dg_search %>%
  ggplot(aes(x="", y=proportion, fill=decision_on_giving_search)) +
  geom_bar(width=1, stat = "identity", color = "white") +
  coord_polar("y", start = 0) +
  geom_text(aes(label=paste(proportion, '%')), size=4, position=position_stack(vjust = 0.5)) +
  theme_void() +
  labs(fill="Decision search") +
  theme(plot.title=element_text(face="bold"))

dg_search_detail <-
  dg %>%
  filter(decision_on_giving_search=="Yes" | decision_on_giving_search=="Sometimes") %>%
  select(starts_with('decision_on_giving_search'))

dg_search_detail <-
  dg_search_detail %>%
  select(-decision_on_giving_search,
         -decision_on_giving_search_consult_other_source)

dg_search_df<-data.frame('Method'=c(''),
                         'No'=c(''),
                         'Yes'=c(''),
                         'NA'=c(''))

dg_search_lst<-c('Read information provided by the charity',
                 'Contact the charity directly',
                 'Consult the CRA website',
                 'Consult the website of a non-profit or private organization',
                 'A general online search',
                 'Ask someone',
                 'Other'
)

i <- 1
for (r in colnames(dg_search_detail)){
  counted<-
    dg_search_detail %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  
  dg_search_df<-rbind(dg_search_df, c(dg_search_lst[i], counted$proportion))
  i<-i+1
}

dg_search_df<-dg_search_df[-c(1),]

dg_search_df %>% 
  as_tibble() %>%
  relocate(Yes, .after=Method) %>%
  kable(caption = "Decision On Giving search")
```

It shows that people usually 

```{r reasonsforgiving, echo=FALSE}
# choose the data who has donated in the past 12 months
reasons_for_giving <-
  gss %>%
    select(reas_giv_pers_affected,
           reas_giv_tax_cred,
           reas_giv_religious,
           reas_giv_spiritual_beliefs,
           reas_giv_cause,
           reas_giv_compassion,
           reas_giv_community_contribution,
           reas_giv_being_asked)
  # filter(reas_giv_pers_affected != "NA")

rg_lst<-c('Personally affected',
          'Tax Credit',
          'Religious',
          'Spiritual beliefs',
          'Cause',
          'Compassion',
          'Community contribution',
          'Being asked')

rg<-data.frame('Reason'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(reasons_for_giving)){
  counted<-
    reasons_for_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  rg<-rbind(rg, c(rg_lst[i], counted$proportion))
  i<-i+1
}
rg<-rg[-c(1),]

rg %>% 
  as_tibble() %>%
  relocate(Yes, .after=Reason) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Reasons for giving")
```

```{r reasonsfornotgiving, echo=FALSE}
reasons_for_not_giving <-
  gss %>%
    select(starts_with("reas_not_giv")) %>%
  mutate(reas_not_giv_charity_fraud = 
           case_when(reas_not_giv_charity_fraud == "Agree" ~ "Yes",
                     reas_not_giv_charity_fraud == "Disagree" ~ "No",
                     reas_not_giv_charity_fraud == "NA" ~ "NA"),
         reas_not_giv_so_many_org = 
           case_when(reas_not_giv_so_many_org == "Agree" ~ "Yes",
                     reas_not_giv_so_many_org == "Disagree" ~ "No",
                     reas_not_giv_so_many_org == "NA" ~ "NA"))

ng_lst<-c('Already gave',
          'Cannot afford',
          'Not being asked',
          'Did not know where to donate',
          'Hard to find cause',
          'Volunteered instead',
          'Gave directly to people',
          'Tax credit not incentive enough',
          'Money would not be used efficient',
          'Not efficient fundraising',
          'Not efficient impact',
          'Not efficient explanation',
          'Not efficient - other',
          'Did not like the way requested',
          'Request time of the day',
          'Request number',
          'Request tone',
          'Request multiple times',
          'Request method',
          'Request - other',
          'So many organizations',
          'Charity fraud'
          )

ng<-data.frame('Reason'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(reasons_for_not_giving)){
  counted<-
    reasons_for_not_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  ng<-rbind(ng, c(ng_lst[i], counted$proportion))
  i<-i+1
}
ng<-ng[-c(1),]

ng %>% 
  as_tibble() %>%
  relocate(Yes, .after=Reason) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Reasons for not giving")
```

```{r othergiving, echo=FALSE}
other_giving<-
  gss %>%
  select(starts_with("other_giv_"))

og_lst<-c('Food bank',
          'Clothings, toys, or household',
          'Bequest in reespond will',
          'Direct to people',
          'Direct to stranger',
          'Direct to personal crowdfunding',
          'Direct to family outside household',
          'Direct to family outside household in Canada',
          'Direct to family outside household outside Canada',
          'Direct other')

og<-data.frame('Other_giving'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(other_giving)){
  counted<-
    other_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  og<-rbind(og, c(og_lst[i], counted$proportion))
  i<-i+1
}
og<-og[-c(1),]

og %>% 
  as_tibble() %>%
  relocate(Yes, .after=Other_giving) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Other givings")
```


```{r donationresponse, fig.cap = "Donations made by responding to methods", echo = FALSE}
# Graph showing the counts for y/n for financial giving

# select the required variables
# rename the variables
financial_giving <- gss %>% 
  select(starts_with("fin_giving_"), caseid) %>% 
  rename("mail" = "fin_giving_mail",
         "telephone" = "fin_giving_telephone",
         "tv" = "fin_giving_tv",
         "online" = "fin_giving_online",
         "own initiative" = "fin_giving_own_initiative",
         "charity event" = "fin_giving_charity_event",
         "in mem of someone" = "fin_giving_in_mem_of_someone",
         "work" = "fin_giving_work",
         "door to door canvassing" = "fin_giving_door_to_door_canvassing",
         "shopping centre" = "fin_giving_shopping_centre",
         "place of worship" = "fin_giving_place_of_worship",
         "sponsoring" = "fin_giving_sponsoring",
         "other" = "fin_giving_other")

# tidy data
financial_giving <- financial_giving %>% 
  pivot_longer(!caseid,
               names_to = "name",
               values_to = "response")

# plot data
graph_1 <-
  financial_giving %>% 
  ggplot(aes(x = name, fill = response)) +
  geom_bar(position = "dodge") +
  labs(x = "Methods responded to",
       y = "Number of survey respondents") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(palette = "PuBu")

graph_1 + theme(legend.position = "bottom")
```


```{r numberdonations, echo = FALSE}
# total number of financial donations
# obtain the total number of financial donations from the raw dataset
total_number_fin_donation <- raw_data %>% 
  select(gs1dntot)

# rename column with a more appropriate name
clean_total_fin_donation <- total_number_fin_donation %>% 
  rename(number_of_financial_donations = gs1dntot)

# obtain the data for those who did not skip the question
clean_total_fin_donation_1 <- 
  clean_total_fin_donation %>% 
  filter(number_of_financial_donations != 96)

# number of people who skipped this question
skipped_response <-
  clean_total_fin_donation %>% 
  filter(number_of_financial_donations == '96') %>% 
  count()

# distribution of total number of financial donations made in the past 12 months
clean_total_fin_donation_1 %>% 
  ggplot(aes(x = number_of_financial_donations)) +
  geom_histogram(bins = 40, fill = "lightblue") +
  labs(x = "Total number of financial donations",
       y = "Number of donations") +
  theme_minimal()
```

```{r}
# investigate the total number of donations made through different means

```



```{r totalamount, echo = FALSE}
# total amount of donations
total_amount <- gss %>% 
  select(totl_amount_don12) %>% 
  rename(total = totl_amount_don12)

# data for only those who responded
total_amount_1 <- total_amount %>% 
  filter(total != 999999999.96)

# number of valid skipped responses
valid_skip_responses <- total_amount %>% 
  filter(total == 999999999.96) %>% 
  count()

# distribution of total amount of donations
total_amount_1 %>% 
  ggplot(aes(x = total)) +
  geom_histogram(bins = 30, fill = "lightblue") +
  labs(x = "Total amount of donations",
       y = "Number of donations") +
  theme_minimal()

```

```{r income, echo = FALSE}
# obtain income data
# rename the variables name
income_group <- raw_data %>% 
  select(incg2, famincg2) %>% 
  rename(personal_income_group = incg2,
         family_income_group = famincg2)

# recode the variable values
income_group <- income_group %>% 
  mutate(personal_income_group = 
           recode(personal_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more"),
         family_income_group = recode(
           family_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more")
           )

# distribution of income
income_group %>% 
  ggplot(aes(x = personal_income_group)) +
  geom_bar(fill = "lightblue") +
  labs(x = "Personal Income group before taxes",
       y = "Number of respondents") +
  theme_minimal() +
  coord_flip()

income_group %>% 
  ggplot(aes(x = family_income_group)) +
  geom_bar(fill = "orange") +
  labs(x = "Family Income group before taxes",
       y = "Number of respondents") +
  theme_minimal() +
  coord_flip()
```


# extra questionnaires
- How much do you donate proportional to your income? (I donate 10% of my income every year, etc.)
