
---
title: "Output"
author: 
  - Pascal Lee Slew
  - Yunkyung Park
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Giving is one way to get involve with the community. Statistics Canada has been collecting data on giving, volunteering,and participating since 1997. In this paper, we focused on exploring the giving aspect, especially the reasons for giving, the decisions on giving and the behavioral patterns on giving. We found that about 68% of respondents read information provided by charity before making a decision. About 37% of respondents responded positively to the all the reasons for giving the survey questionnaire and people in higher income bracket tend to donate more. The analysis and monitoring of these social trends give a reflection of the living conditions of the Canadian public and has significant implications for public policy and program decisions."
output:
  bookdown::pdf_document2
toc: FALSE
#bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

Charities form an important part of the Canadian social fabric. The charitable sector plays a crucial role in the Canadian economy, providing essential services to the vulnerable. The services provided by charities bring multitudes of societal benefits in a wide array of sectors. Charitable giving is no longer what it was in previous years. We are seeing a decline in charity giving. Canadians are giving less. It is becoming a cause for concern since a large number of Canadians rely on the services provided by those charities to live.

Statistics Canada has collected data on giving, volunteering and participating since 1997 at an interval of about 3 years. The General Social Survey on Giving, Volunteering and Participating (GSS GVP) collects data data through different questionnaires and the objective is to observe any changes in societal behaviour and pattern through analysis of the data collected, providing vital information to governmental bodies. 

This paper focuses on the giving aspect of the 2018 GSS GVP. It gives a descriptive analysis on giving behaviour of Canadians. We have looked at the reasons why Canadians donate, the decisions on giving and reasons for not giving more among other aspects to analyze the giving behaviour of Canadian and provide some insights on how that may help fundraisers in the future.

An interesting point found was that respondents earning 100000 but less than 125000 did not donate as much compared to other income groups. We also found that most respondents made between 1 to 5 donations in the past 12 months, indicating that they do not donate on a regular basis. 

The rest of this paper will be as follows: 


# Data

## Data Collection

The data was obtained from Canadian 2018 General Social Survey (GSS) on Giving, Volunteering, and Participation (GVP). The survey was conducted by interviewing individuals 15 years of age or over in Canada's ten provinces excluding Yukon, Northwest Territories, and Nunavut. It was collected from September 4 to December 28, 2018 by Statistics Canada. Respondents were given two option to complete the questionnaire online and via the telephone. 


## Survey Methodology

How respondents were selected?
It applied the age-order selection method. It is important to note that all respondents were not contacted the same way; some were contacted through mail, others by telephone.

Changes to survey questionnaire
content was updated from 2004 to 2018 using experience from prior surveys, notably clarity of questions and consultations with stakeholders from charity organizations. 

Another important feature of the 2018 GSS GVP is that not all respondents have the same number of questions. For example, only respondents who donated at least 10 dollars were  asked questions about decisions and reasons for giving. And those who donated more than 1150 dollars did not have to answer why they did not choose to donate more.

For the online questionnaire, a pilot test was conducted online and by telephone from February 2018 to March 30 2018.

sample frame
it has not been consistent over the years. In 2004, 2007 and 2010, random digit dialling was used. From 2013, the sample frame included "cell phone only" households, which is a population not covered by RDD. 

Internet option - adapt to technology and demands on Canadians'time. - flexibility and convenience

target population
all persons who are at least 15 years old with the exceptions of residents of the Yukon, Northwest Territories and Nunavut as well as full-time residents of institutions

proceeded with stratification
stratification sampling with the provinces divided into strata and many of the Census Metropolitan Areas (CMA) considered as separate strata. totaling 27 stata

frame
2 parts - lists of telephone numbers in use (both landline and cellular) available to Statistics Canada
- address register for mail code for online option

for each stratum, a siimple random sample without replacement was selected.

rejective sampling - 'rejecting' a portion of population with a given probability to allow more time and effort to be spent trying to find the population of interest, volunteers

target sample size was 20000 but actual number is 16149
minimum sample size for each stratum was determined beforehand to ensure acceptable sample variability

majority of interviewers had previous experience with gss or csgvp

- explain importance of survey and encourage participation

response rate was 41.9%
less than previous gss but not comparable



## Discussion and Results




strength
- large coverage with detailed/ specific questions
- well documented
- 

weaknesses
- too specific, might take too long to respond to all
- some data obtained from other sources like income from CRA 2017 tax filing, consent for survey?
- first time online version, may not be consistent with previous versions
- data quality with online questionnaire
- 

```{r library, echo=FALSE, message=FALSE}
# install.packages('bookdown')
library(dplyr)
library(tidyverse) 
library(janitor)
library(ggplot2)
library(knitr)
library(kableExtra)
library(bookdown)
```

```{r read_data, echo=FALSE, message=FALSE, warning=FALSE}
gss<- read_csv('data1.csv', show_col_types = FALSE)
raw_data <- read_csv('gssv.csv')
```

# I. Introduction

This dataset contains core and 

```{r age, echo=FALSE}
age <-
  gss %>%
  mutate(age_group = 
           case_when(age == "15 to 24 years" ~ "15-24",
                     age == "25 to 34 years" ~ "25-34",
                     age == "35 to 44 years" ~ "35-44",
                     age == "45 to 54 years" ~ "45-54",
                     age == "55 to 64 years" ~ "55-64",
                     age == "65 to 74 years" ~ "65-74",
                     age == "75 years and over" ~ ">=75"
                     ),
         age_group = factor(age_group,
                            levels = c("15-24",
                                       "25-34",
                                       "35-44",
                                       "45-54",
                                       "55-64",
                                       "65-74",
                                       ">=75"
                                       )
                            )
         )
age %>%
  ggplot(mapping=aes(x=age_group, fill=gender)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(x='Age-group of respondent', y='Number of respondents', 
       title='Distribution of age-groups and gender in the 2018 GSS survey',
       fill='Gender')
```
The age distribution of survey respondents is shown in Figure \@ref(fig:age). Among the selected household, one member was chosen to represent the household. It shows that it was more likely that the female in their middle ages was selected to complete the questionnaire. 


```{r maritalstatus, echo=FALSE, fig.cap='Marital Status and Income'}
age %>%
  ggplot(mapping=aes(x=fam_income_quintile)) +
  geom_bar() +
  theme_minimal() +
  labs(x='', title='Marital Status') +
  theme(axis.text.x = element_text(angle=15)) +
  facet_wrap(vars(marital_status),
             dir = "v",
             scales = "free")
```

Marital status is important in tracking the evolution of social attitudes. Figure \@ref(fig:maritalstatus) indicates that the most common family structure in Canada is married couples. When the marital status is combined with the family income, it can be used to indicate the socio-economic status of various types of families. Figure \@ref(fig:maritalandincome) tells that the distribution of the family income quintile is similar between the married couples and the common-law couples, and the remaining tends to have the similar distribution.

```{r donation, echo=FALSE, fig.cap='Whether they have made the donation in the past 12 months'}
giving <-
  gss %>%
    count(giving_flag)
giving %>%
  ggplot(mapping=aes(x=giving_flag, y=n, fill=giving_flag)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  geom_text(aes(label=paste(round(n/sum(n),4)*100, '%')), size=4, position=position_stack(vjust = 0.5)) +
  theme_minimal() +
  labs(x='',y='count') +
  theme(legend.position='none')
```

Figure \@ref(fig:donation) shows that 75.26% of the respondents have answered that they have made any financial donations in the past 12 months.

```{r awareness, echo=FALSE, fig.cap="Awareness of the charity organizations"}
dg <- gss %>% 
  select(starts_with('decision_on_giving_'))

dg_awareness<-
  dg %>%
    select(decision_on_giving_know_how_to_verify, decision_on_giving_org_monitoring)

awareness_lst<-c('Know how to verify whether an organization is a registered charity',
                 'Monitor how charities in Canada use their donations')

df_awareness<-data.frame('Method'=c(''),
                         'No'=c(''),
                         'Yes'=c(''),
                         'NA'=c(''))
i <- 1
for (r in colnames(dg_awareness)){
  counted<-
    dg_awareness %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  
  df_awareness<-rbind(df_awareness, c(awareness_lst[i], counted$proportion))
  i<-i+1
}

df_awareness<-df_awareness[-c(1),]

df_awareness %>% 
  as_tibble() %>%
  relocate(Yes, .after=Method) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Awareness of the charity organizations", format="latex") %>%
  row_spec(0, bold=TRUE) %>%
  kable_styling()
```

Based on Table \@ref(tab:awareness), we can first observe that most people do not know how to check if a charity is registered in Canada. A "registered charity" refers to a charitable organization that is registered with the Canada Revenue Agency (@CRA). The donation made to the registered charity is exempted from paying income tax, and the charity can issue tax receipts for donations it receives (@CRA). Next, most charities openly share their information on how the donations are being used for its accountability and transparency (@charity). However, we can tell that people do not monitor how charities use their donations.

### How to decide where to donate 




```{r dgsearch, echo=FALSE}
dg_search <-
  dg %>% 
  mutate(decision_on_giving_search =
           recode(
             decision_on_giving_search,
             'Yes'='Yes',
             'No'='No',
             'Sometimes'='Sometimes',
             'You never consider donating to a charity that you have no...' = 'Never considered donating to a new charity')) %>%
  count(decision_on_giving_search) %>% 
  mutate(proportion = round(n/sum(n),4)*100)

dg_search %>%
  ggplot(aes(x="", y=proportion, fill=decision_on_giving_search)) +
  geom_bar(width=1, stat = "identity", color = "white") +
  coord_polar("y", start = 0) +
  geom_text(aes(label=paste(proportion, '%')), size=4, position=position_stack(vjust = 0.5)) +
  theme_void() +
  labs(fill="Decision search") +
  theme(plot.title=element_text(face="bold"))

dg_search_detail <-
  dg %>%
  filter(decision_on_giving_search=="Yes" | decision_on_giving_search=="Sometimes") %>%
  select(starts_with('decision_on_giving_search'))

dg_search_detail <-
  dg_search_detail %>%
  select(-decision_on_giving_search,
         -decision_on_giving_search_consult_other_source)

dg_search_df<-data.frame('Method'=c(''),
                         'No'=c(''),
                         'Yes'=c(''),
                         'NA'=c(''))

dg_search_lst<-c('Read information provided by the charity',
                 'Contact the charity directly',
                 'Consult the CRA website',
                 'Consult the website of a non-profit or private organization',
                 'A general online search',
                 'Ask someone',
                 'Other'
)

```


```{r search, echo=FALSE}
i <- 1
for (r in colnames(dg_search_detail)){
  counted<-
    dg_search_detail %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  
  dg_search_df<-rbind(dg_search_df, c(dg_search_lst[i], counted$proportion))
  i<-i+1
}

dg_search_df<-dg_search_df[-c(1),]

dg_search_df %>% 
  as_tibble() %>%
  relocate(Yes, .after=Method) %>%
  kable(caption = "Decision On Giving search")
```

Among those who search before they donate to a charity, Table \@ref(tab:search) shows that people are mostly likely to read information provided by the charity and do a general online search. It should be noted that even though the list of the registered charitities can be viewed from the CRA website, people did not consider consulting the CRA website before making donations. 

Next, we observed whether people do a research on the organizations before making donations, and if they do, what methods are being used to search for the charity. Figure \@ref(fig:dgsearch) shows that if the respondent have considered making donations, then they usually search for information on the charity before giving. 


```{r reasonsforgiving, echo=FALSE}
# choose the data who has donated in the past 12 months
reasons_for_giving <-
  gss %>%
    select(reas_giv_pers_affected,
           reas_giv_tax_cred,
           reas_giv_religious,
           reas_giv_spiritual_beliefs,
           reas_giv_cause,
           reas_giv_compassion,
           reas_giv_community_contribution,
           reas_giv_being_asked)
  # filter(reas_giv_pers_affected != "NA")

rg_lst<-c('Personally affected',
          'Tax Credit',
          'Religious',
          'Spiritual beliefs',
          'Cause',
          'Compassion',
          'Community contribution',
          'Being asked')

rg<-data.frame('Reason'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(reasons_for_giving)){
  counted<-
    reasons_for_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  rg<-rbind(rg, c(rg_lst[i], counted$proportion))
  i<-i+1
}
rg<-rg[-c(1),]

rg %>% 
  as_tibble() %>%
  relocate(Yes, .after=Reason) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Reasons for giving")
```

```{r reasonsfornotgiving, echo=FALSE}
reasons_for_not_giving <-
  gss %>%
    select(starts_with("reas_not_giv")) %>%
  mutate(reas_not_giv_charity_fraud = 
           case_when(reas_not_giv_charity_fraud == "Agree" ~ "Yes",
                     reas_not_giv_charity_fraud == "Disagree" ~ "No",
                     reas_not_giv_charity_fraud == "NA" ~ "NA"),
         reas_not_giv_so_many_org = 
           case_when(reas_not_giv_so_many_org == "Agree" ~ "Yes",
                     reas_not_giv_so_many_org == "Disagree" ~ "No",
                     reas_not_giv_so_many_org == "NA" ~ "NA"))

ng_lst<-c('Already gave',
          'Cannot afford',
          'Not being asked',
          'Did not know where to donate',
          'Hard to find cause',
          'Volunteered instead',
          'Gave directly to people',
          'Tax credit not incentive enough',
          'Money would not be used efficient',
          'Not efficient fundraising',
          'Not efficient impact',
          'Not efficient explanation',
          'Not efficient - other',
          'Did not like the way requested',
          'Request time of the day',
          'Request number',
          'Request tone',
          'Request multiple times',
          'Request method',
          'Request - other',
          'So many organizations',
          'Charity fraud'
          )

ng<-data.frame('Reason'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(reasons_for_not_giving)){
  counted<-
    reasons_for_not_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  ng<-rbind(ng, c(ng_lst[i], counted$proportion))
  i<-i+1
}
ng<-ng[-c(1),]

ng %>% 
  as_tibble() %>%
  relocate(Yes, .after=Reason) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Reasons for not giving")
```

Table \@ref(tab:reasonsforgiving) demonstrates the reasons for giving. Note that only the respondents who have made the donations in the past 12 months answered this questionnaire. Table \@ref(tab:reasonsforgiving) indicates that it varies for each people why they decide to make donations. Although it turned out to be approximately equal among every reason listed in the survey, the top 2 were spiritual beliefs and religious reasons, which imply that one's faith is one of the biggest reasons why people decide to give money to the charity.


```{r othergiving, echo=FALSE}
other_giving<-
  gss %>%
  select(starts_with("other_giv_"))

og_lst<-c('Food bank',
          'Clothings, toys, or household',
          'Bequest in reespond will',
          'Direct to people',
          'Direct to stranger',
          'Direct to personal crowdfunding',
          'Direct to family outside household',
          'Direct to family outside household in Canada',
          'Direct to family outside household outside Canada',
          'Direct other')

og<-data.frame('Other_giving'=c(''),
               'NA'=c(''),
               'No'=c(''),
               'Yes'=c(''))

i <- 1
for (r in colnames(other_giving)){
  counted<-
    other_giving %>%
    count(.data[[r]]) %>%
    mutate(proportion = round(n/sum(n),4)*100)
  og<-rbind(og, c(og_lst[i], counted$proportion))
  i<-i+1
}
og<-og[-c(1),]

og %>% 
  as_tibble() %>%
  relocate(Yes, .after=Other_giving) %>%
  relocate(No, .after=Yes) %>%
  kable(caption = "Other givings")
```

Next, we observed for the reasons for not giving. If the respondents have made the donations in the past 12 months, then it asked for the reasons why they did not donate more. Table \@ref(tab:reasonsfornotgiving) demonstrates that among the respondents who have donated in the past 12 months, 31.17% of them think that they have already donated enough. 
When the respondents did not donate in the past 12 months, it asked for the reasons for not giving. Table \@ref(tab:reasonsfornotgiving) shows that 42.57% of the respondents decided to volunteer instead of giving. 42.41% of the respondents answered that they did not like the way they were being requested. Additionally, it is worth mentioning that although there are so many organizations, it was not the main reason why people decided not to give money. It was that they did not know where to make donations among those organizations, and they were afraid that money would not be used efficient. Furthermore, 42.32% of the respondents said that they gave money directly to people who were in need of financial assistance.

### Other givings

```{r donationresponse, fig.cap = "Comparing the responses to different methods", echo = FALSE}
# Graph showing the counts for y/n for financial giving

# select the required variables
# rename the variables
financial_giving <- gss %>% 
  select(starts_with("fin_giving_"), caseid) %>% 
  rename("mail" = "fin_giving_mail",
         "telephone" = "fin_giving_telephone",
         "tv" = "fin_giving_tv",
         "online" = "fin_giving_online",
         "own initiative" = "fin_giving_own_initiative",
         "charity event" = "fin_giving_charity_event",
         "in mem of someone" = "fin_giving_in_mem_of_someone",
         "work" = "fin_giving_work",
         "door to door canvassing" = "fin_giving_door_to_door_canvassing",
         "shopping centre" = "fin_giving_shopping_centre",
         "place of worship" = "fin_giving_place_of_worship",
         "sponsoring" = "fin_giving_sponsoring",
         "other" = "fin_giving_other")

# tidy data
financial_giving <- financial_giving %>% 
  pivot_longer(!caseid,
               names_to = "name",
               values_to = "response")

# plot data
graph_1 <-
  financial_giving %>% 
  ggplot(aes(x = name, fill = response)) +
  geom_bar(position = "dodge") +
  labs(x = "Methods responded to",
       y = "Number of survey respondents") +
  coord_flip() +
  theme_minimal() +
  scale_fill_brewer(palette = "PuBu")

graph_1 + theme(legend.position = "bottom")
```

Figure \@ref(fig:donationresponse) shows the number of responses of the different methods responded to. For example, online means the number of people who made donations by responding to an online request which can be an email or social media. From the figure, it is clear that people tend to positively respond to request in shopping centre while other ways of requests received the least number of donations. However, the number of donations does not necessarily mean more amount donated. 

```{r numberdonations, echo=FALSE, fig.cap = "Distribution of the total number of financial donations in the past 12 months"}
# total number of financial donations
# obtain the total number of financial donations from the raw dataset
total_number_fin_donation <- raw_data %>% 
  select(gs1dntot)

# rename column with a more appropriate name
clean_total_fin_donation <- total_number_fin_donation %>% 
  rename(number_of_financial_donations = gs1dntot)

# obtain the data for those who did not skip the question
clean_total_fin_donation_1 <- 
  clean_total_fin_donation %>% 
  filter(number_of_financial_donations != 96)

# number of people who skipped this question
skipped_response <-
  clean_total_fin_donation %>% 
  filter(number_of_financial_donations == '96') %>% 
  count()

# distribution of total number of financial donations made in the past 12 months
clean_total_fin_donation_1 %>% 
  ggplot(aes(x = number_of_financial_donations)) +
  geom_histogram(bins = 40, fill = "lightblue") +
  labs(x = "Total number of financial donations",
       y = "Number of donations") +
  theme_minimal()
```

Figure \@ref(fig:numberdonations) shows the distribution of the number financial donations made by the survey respondents. It is important to note that out of the 16149 respondents, 3995 of them were not asked this question since they did not make any financial donations and this number was removed before making the graph. The most number of financial donations made was 72 but with most other respondents making between 1 and 5 donations in the last 12 months.




```{r total, fig.cap = "Distribution of total amount donated in the past 12 months", echo = FALSE}
# total amount of donations
total_amount <- gss %>% 
  select(totl_amount_don12) %>% 
  rename(total = totl_amount_don12)

# data for only those who responded
total_amount_1 <- total_amount %>% 
  filter(total != 999999999.96)

# number of valid skipped responses
valid_skip_responses <- total_amount %>% 
  filter(total == 999999999.96) %>% 
  count()

# distribution of total amount of donations
total_amount_1 %>% 
  ggplot(aes(x = total)) +
  geom_histogram(bins = 30, fill = "lightblue") +
  labs(x = "Total amount of donations",
       y = "Number of donations") +
  theme_minimal()

# total amount donated
total_sum <- sum(total_amount_1$total)

g <-total_amount_1 %>% 
  filter(total > 20000)
```

Figure \@ref(fig:total) illustrates how the donation amounts are distributed with 0.5 dollars as the least amount donated and $60,000 as the largest amount donated in the past 12 months. 3995 respondents chose to skip this question. A total of 9.3 million dollars have been donated by the 12154 Canadian respondents. From figure \@ref(fig:total), most respondents donated amounts less than 200 dollars while only 28 out of the 12154 respondents donated more than 20000 dollars.


```{r income, fig.cap = "Comparing the number of respondents in the different income groups", echo = FALSE}
# obtain income data
# rename the variables name
income_group <- raw_data %>% 
  select(incg2, famincg2) %>% 
  rename(personal_income_group = incg2,
         family_income_group = famincg2)

# recode the variable values
income_group <- income_group %>% 
  mutate(personal_income_group = 
           recode(personal_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more"),
         family_income_group = recode(
           family_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more")
           )

# distribution of income
income_group %>% 
  ggplot(aes(x = personal_income_group)) +
  geom_bar(fill = "lightblue") +
  labs(x = "Personal Income group before taxes",
       y = "Number of respondents") +
  theme_minimal() +
  coord_flip()


```

Figure \@ref(fig:income) shows the distribution of personal income groups of the survey respondents. The data was obtained using the CRA income tax filing in 2017. We can see that most respondents were from low to middle income background. The highest proportion of respondents were found in the 25000 to 49999 income group (29.95%) and 78.23% of the respondents earned less than $75000 annually before taxes.




```{r averagedonate, echo=FALSE}
# age income 
donate_income_group <- raw_data %>% 
  select(incg2, gs1datot) %>% 
  rename(personal_income_group = incg2,
         total_amount_donated = gs1datot)

donate_income_group <- donate_income_group %>% 
  mutate(personal_income_group = 
           recode(personal_income_group,
                  "1" = "less than 25000",
                  "2" = "25000 to 49999",
                  "3" = "50000 to 74999",
                  "4" = "75000 to 99999",
                  "5" = "10000 to 124999",
                  "6" = "125000 or more")
           )

# create summary table

donate_income_group %>% 
  filter(total_amount_donated != 999999999.96) %>% 
  group_by(personal_income_group) %>% 
  summarise(mean = round(mean(total_amount_donated), 2),
            sd = round(sd(total_amount_donated), 2)) %>% 
  kable(caption = "Summary statistics of total amount donated per personal income group",
        col.names = c("Personal Income group", "Average amount donated", "Standard deviation of amount donated"),
        linesep = "",
        align = c("c", "c", "c"))

```

Table \@ref(tab:averagedonate) shows the mean of the total amount donated for each personal income group as well as its variability. We can note that in general, respondents in higher income group tends have donated more on average which is not surprising. We can also observed that the mean amount donated for the 75000 to 99999 and 100000 to 124999 income group do not differ by much ($63.19) but the variability is much larger which can explain why the difference is so small. There must be some respondents who donated much larger amount compared to what others in the same income group have donated.

```{r, echo = FALSE}
# age and donation relation
donate_age_group <- raw_data %>% 
  select(agegr10, gs1datot) %>% 
  rename(age_group = agegr10,
         total_amount_donated = gs1datot)

donate_age_group <- donate_age_group %>% 
  mutate(age_group = recode(age_group,
                            "1" = "15-24",
                            "2" = "25-34",
                            "3" = "35-44",
                            "4" = "45-54",
                            "5" = "55-64",
                            "6" = "65-74",
                            "7" =  ">=75")
           )

# create summary table

donate_age_group %>% 
  filter(total_amount_donated != 999999999.96) %>% 
  group_by(age_group) %>% 
  summarise(mean = round(mean(total_amount_donated), 2),
            sd = round(sd(total_amount_donated), 2)) %>% 
  kable(caption = "Summary statistics of total amount donated per age group",
        col.names = c("Age group", "Average amount donated", "Standard deviation of amount donated"),
        linesep = "",
        align = c("c", "c", "c"))

```

## Awareness of the charitable organizations

A "registered charity" refers to a charitable organization that is registered with the Canada Revenue Agency (@registeredcharity). There is a difference between a registered charity and a non-profit organization. Although both registered charities and non-profit organizations run on a non-profit basis, the registered charity should have more than one of the following charitable purposes: the relief of poverty, the advancement of education, the advancement of religion, or other purposes that benefit the community (@registeredcharity). On the other hand, the non-profit organization operates exclusively for social welfare, civic improvement, pleasure, recreation, or any other purpose except profit (@registeredcharity). Recreational groups or certain amateur sports organizations are regarded as non-profit organizations (@registeredcharity). 

If a charity is registered, then the group can issue official donation receipts for income tax purposes to donors and is exempt from paying income tax (@registeredcharity). In addition, donors can receive income tax credits for the financial donations that they have made (@registeredcharity). Therefore, it is often the case where people decide to make donations for the income tax credit, which can be observed in Table \@ref(tab:reasonsforgiving). However, Table \@ref(tab:awareness) tells that most people do not know how to check whether a charity is registered under the Income Tax Act. Furthermore, a list of the registered charities can be viewed from the CRA website, but Table \@ref(tab:search) demonstrates that people do not consider consulting the CRA website before making a decision on where to make donations. Meanwhile, Table \@ref(tab:reasonsfornotgiving) tells that 33.07% of the respondents answered that tax credit is not incentive enough. This can be asked in the supplementary survey to figure out what changes can be made to let people be aware of the registered charity.

Next, donors need to ensure that the charity is trustworthy. Most charities, especially when it is a good charity, openly share their information on how the donations are being used for its accountability and transparency (@charity). They post financial statements on their websites to ensure how the donations are being used (@CPA). However, Table \@ref(tab:awareness) indicates that most people do not monitor how charities use their donations. While 68.64% of the respondents answered that they do research about the charity before making donations, they do not keep a check on how their money is being used. When we invest in stocks or mutual funds, we keep checking their price so that we are not losing any money. This is the same for the donation. We should monitor the charity constantly to make sure that the money is not being wasted. 

## Trend

Figure \@fig(ref:donation) shows that 75.26% of the respondents have made the financial donation in the last 12 months. Among the 16,149 respondents, 3,995 people did not give money to the charitable or non-profit organizations, which constituted to a 24.74% of the respondents. Figure \@ref(fig:totalamount) shows that there are few who made a large amount of donations while most donors donated less than \$200. Similar distribution can be observed in Figure \@ref(fig:numberdonations) where it illustrates the distribution of the total number of financial donations made in the past 12 months. There were few people who made donations frequently while most donors made donations between less than 5 times. 

## Limitations and Weaknesses

First of all, households without telephones were excluded from the survey population. It was the seventh time that the GVP was run, and in the previous GVP surveys, only paper questionnaires and telephone were used to collect data. However, in the survey conducted in 2018, a paper questionnaire was replaced by an online. Since the paper option is no longer available, it was crucial to include the "cell phone only" households to cover more Canadian population. However, the survey missed the households without the Internet and cell phone, and this population is harder to reach. Since most people have access to thhe Internet and have their own cell phones, we can assume that they are the ones who are likely in need of financial help. This might have changed our result in Table \@ref(tab:reasonsfornotgiving), which is about the reasons for not giving. 

Second, this survey covers a wide range of giving with detailed questions. However, it is too detailed at the same point that it ended up containing so many questionnaires in the survey. Therefore, there are so many missing values in many variables. This can be observed in multiple figures and tables in the paper. For example, in Table \@ref(tab:reasonsfornotgiving), most entries have more than half of the missing values. Also, as there were so many questions in the survey, it is probable that the respondents did not spend enough time completing the questionnaires, especially if it was conducted online. For example, in Table \@ref(tab:reasonsforgiving) and Table \@ref(tab:reasonsfornotgiving), we can observe that the proportion of the respondents who answered yes is similar among most of the items. Since these questions were asked one after another, it might be the case that the respondents did not read the questions thoroughly before answering them. 

Third, the survey was conducted from September to December in 2018. It asked for the financial donation given to a charitable or other non-profit organization during the 12-month reference period preceding the survey. Since every respondent completed the questionnaire at a different time, it is hard to define the exact time frame that was used in collecting the data. It would have been better if the period was specified such as the donations made during the period of January 1st to December 31st in 2017. In addition, even though the respondents are most likely to answer the questionnaires based on their donation history in 2018, there are several questions which asked on an annual basis. For example, the data for personal income was obtained using the CRA income tax filing in 2017. Thus, the time period does not match among the questions. Furthermore, it is not mentioned whether the respondents have consented to use their income tax information. 

# Appendix

## Supplementary survey



# extra questionnaires
- How much do you donate proportional to your income? (I donate 10% of my income every year, etc.)
- Are you thinking of making any donations in the near future?
- What changes will make you to want to donate in the near future?
- 
- more categories for gender, with possibly a space to specify like other, specify
- when people feel like donating
- motivation for donating, in the family
- will they consider donating in the future
- 


\newpage

# Appendix

## Supplementary survey



# extra questionnaires
- How much do you donate proportional to your income? (I donate 10% of my income every year, etc.)

- more categories for gender, with possibly a space to specify like other, specify
- when people feel like donating
- motivation for donating, in the family
- will they consider donating in the future
- 
